Bayesian model in JAGS

### 2023-01-30 4_RSV_disease_burden/4_RSV_disease_burden.Rmd
update

---
title: "Bayesian model in JAGS"
subtitle: "Tutorial 3 for transition"
author: "Zhe Zheng (Gigi)^[zhe.zheng@yale.edu]"
date: "2023-01-16"
fontsize: 10pt
output: 
  beamer_presentation:
  theme: CambridgeUS
  colortheme: "beaver"
  fonttheme: "structurebold"
---

@@

- Section 1: Background knowledge of estimating respiratory virus attributable disease burden (hospitalizations/mortality etc) with statistical models.
\
\
- Section 2: Background knowledge of Hierarchical Bayesian regression.
\
\
- Section 3: Background knowledge of using Hierarchical Bayesian regression to estimate RSV attributable hospitalizations.
\
\
- Section 4: R code for running the Hierarchical Bayesian regression to estimate RSV attributable hospitalizations in older adults by age and risk groups.

@@

## Section 2: Background knowledge of Hierarchical Bayesian regression

>>>>

---
title: "Bayesian model in JAGS"
subtitle: "Tutorial 3 for transition"
author: "Zhe Zheng (Gigi)^[zhe.zheng@yale.edu]"
date: "2023-01-16"
fontsize: 10pt
output: 
  beamer_presentation:
  theme: CambridgeUS
  colortheme: "beaver"
  fonttheme: "structurebold"
header-includes:
  - \usepackage{pdfpages}
  - \setbeamercolor{background canvas}{bg=}
  - \makeatletter\beamer@ignorenonframefalse\makeatother
---

@@

- Section 1: Background knowledge of estimating respiratory virus attributable disease burden (hospitalizations/mortality etc) with statistical models.
\
\
- Section 2: Background knowledge of Hierarchical Bayesian regression
\
\
- Section 3: Hierarchical Bayesian regression to estimate RSV attributable hospitalizations in older adults by age and risk groups and its R code.

@@

## Section 2: Background knowledge of using Hierarchical Bayesian regression to estimate RSV attributable hospitalizations
The benefits of using a Hierarchical Bayesian structure is that it can shrink the uncertainty of inference while allow between-group variability. 
\
\
For example, if we would like to estimate the baseline hospitalizations $\beta_{0jk}$ in age $j$ and socioeconomic group $k$, we can treat this parameter as (1) independent, meaning that the baseline hospitalizations are different in each group and we will need to estimate 27 parameters $\beta_{0j=1,k=1}$, $\beta_{0j=1,k=2}$...$\beta_{0j=9,k=3}$, (2) the same, meaning that the baseline hospitalizations are the same in all age and socioeconomic group and we will  estimate one parameter $\beta_0$, and (3) random draws from the a normal distribution, meaning that baseline hospitalizations are normally distributed in the whole population and baseline hospitalization of each subpopulation $jk$ is an realization of the distribution (a point in the distribution). In this case, we will estimate the mean and variations of the normal distribution and then get the subpopulation estimates.

## Background knowledge of Hierarchical Bayesian regression: Benefits

If we model the parameters in Bayesian framework, the third scenario corresponds to a Hierarchical Bayesian model.
\
\
Compared with treating the parameters as group-specific, Bayesian Hierarchical structure pools information across groups to help shrink the uncertainty of the single group estimates. 
\
\
Compared with treating the group specific parameters as homogeneous, Bayesian Hierarchical structure gives the flexibility and possibility of estimating group specific parameters. It can help evaluate the hypothesis of between-group variation.

## Visualization of a Hierarchical Bayesian structure:

![Hierarchical Bayesian structure](./Hier_Bay_figure.png)

```{r,eval=FALSE,echo=FALSE}
curve(dnorm(x, 0, 100), from=-200, to=200,main="prior distribution of mean of beta0",xlab="mu",ylab="dnorm(0,100)")
curve(dinvgamma(x, shape=0.01, rate = 0.01, log = FALSE), from=0, to=200,ylab="dinvgamma(0.01,0,01)",main="prior distribution of\nvariance sigma square",xlab="sigma square")
```

## Further readings on Hierarchical Bayesian regression

- [Bayesian hierarchical modeling by wikipedia](https://en.wikipedia.org/wiki/Bayesian_hierarchical_modeling)

- [Chapter 10 Bayesian Hierarchical Modeling by Drs. Jim Albert and Jingchen Hu](https://bayesball.github.io/BOOK/bayesian-hierarchical-modeling.html)

- [Hierarchical Bayesian Modeling by Dr. Angie Wolfgang](https://astrostatistics.psu.edu/RLectures/hierarchical.pdf)

## Section 3: Hierarchical Bayesian regression to estimate RSV attributable hospitalizations in older adults by age and risk groups and its R code.

\includepdf[pages=2]{Hier_Bay_figure.pdf}

## Hierarchical Bayesian regression structure

\includepdf[pages=3]{Hier_Bay_figure.pdf}

``` {=latex}
\end{frame}
\includepdf[pages=1-7]{model_structure.pdf}
\begin{frame}
```

## R code (JAGS code) for Hierarchical Bayesian regression structure
